### 2014-02-27 paper/concrete.tex
move proofs.

Instead of using a configuration of the form
|iconf iS (fullconf id1 il1 tS1 ie1, fullconf id2 il2 tS2 ie2 ldots)|
as is done in |specLangML alpha|
where every task has its
own ML heap, we would like a single global heap as in
|oneheapiconf iS tS (oneheapfullconf id1 il1 ie1, oneheapfullconf id2 il2 ie2, ldots)|.
Furthermore, we keep up a fictional separation between the portions
of the global heap that tasks can use by forbidding references to
cross task boundaries. Most
of the operational rules can easily be adapted to work with these
new configurations.  In the abstract system, the \textsc{I-fork}
rule created a new heap for the new task to execute in.  Since we
want to forbid references to cross task boundaries, we do not modify
the global heap and instead add an additional premise that the forked
expression |ie| must not contain any references.
Finally, we add the same condition to the \textsc{I-send} rule.
Tasks that try to fork or send an expression that contains a
reference will now get stuck, and will stop executing by the
\textsc{I-noStep} rule.
We call this family of languages |concreteLangMl alpha|, which
is still parametrized by the scheduling policy |alpha|.

@@

As we have outlined the IFC-Mini-ML language with a single heap, we
effectively restrict the abstract semantics in some cases (namely
to not allow the passing of references) as well as using a different
way to represent the heap.  How can we be sure that these
new, \textit{concrete semantics} for our system still satisfy
termination-sensitive non-interference?

Intuitively, there are two scenarios in which non-interference
can be violated when restricting the transition relation:
\begin{itemize}
  \item Adding an additional premise to one of the rules that inspects
  information which is above the security label of the task currently
  executing.  This would allow another task to observe the termination
  of the other if both are at the same security level, and thus observe
  the secret information that was used in the additional
  premise of the rule.
  \item Adding additional restrictions to transition rules can
  also cause the whole configuration to get stuck, which can leak
  information through the termination channel.
\end{itemize}

Next, we will make these two intuitions precise and give a general
definition for how non-interference can be preserved.  Then, we
show that the definition in Figure~\Red{XXX} satisfy these
conditions.

@@

Now, we can propose the following two-step process to ensure that
TSNI is preserved:
\Red{We should find a better name for this, or not use a name, or something.}
\begin{definition}[Non-Interference preserving, restricted isomorphism (NPRI)]
  We call an IFC language
  |concreteLang alpha targetLang| (which can be thought of as
  the concrete language) a \textit{non-intereference preserving,
  restricted isomorphism} with regard to the IFC language
  |specLang alpha targetLang|
  (which can be thought of as the abstract language) if the following
  two conditions hold:
  \begin{enumerate}
    \item Any of the reduction rules of the abstract
    semantics |specLang alpha targetLang|
    presented in Section~\ref{sec:retrofit} can
    be restricted by adding a predicate $P$ to the premise of
    any rule other than \textsc{I-noStep}.  The predicate $P$
    can depend only on the \textit{erased} configuration
    |erase il ic|, where |il| is the label of the first task
    in the task list.
    \item The concrete language
    |concreteLang alpha targetLang|
    must be isomorphic to the abstract semantics |specLang alpha targetLang|
    with the additional predicates $P$ on the rules.
  \end{enumerate}
\end{definition}

\begin{theorem}[NPRI implies TSNI]
  If a language |concreteLang alpha targetLang| is a NPRI, then
  |concreteLang alpha targetLang| satisfies TSNI.
\end{theorem}

With this machinery in place, we can now state and prove that our
single-heap ML-like IFC language is NPRI, and thus non-interfering.

@@

\begin{proof}
  It is easy to see that condition 1.\ of NPRI is satisfied, since
  we only add a side condition to the rules \textsc{I-fork} and
  \textsc{I-send}, and the condition only inspects the expression
  of the current task, which is by the definition of |erasef il| not
  erased.

@@

\Red{At some point in this section we should say a little about
  why the abstract setting needs complete isolation (because it
  works for \textit{any} language), and that this is sometimes
  what you want.}

@@

Isomorphism requires that |finv| is functorial over reduction relations
as well, i.e. that we can extract a concrete reduction sequence from an abstract one.

\begin{definition}[Isomorphism of information-flow control languages]
    A language \con{|(C, .->, erasef l)|} is
    isomorphic to a language \abs{|(iC, .->, erasef l)|} if there exist total functions
    |f : tC -> iC| and |finv : iC -> tC| such that |f .. finv = idf tC|, |finv .. f = idf iC|, and both |f| and |finv| are
    functorial over both $l$-equivalences and |.->|.
%(if |cc clequiv
%   cc'| then
%   |(ap f cc) alequiv (ap f cc')| and if |ac alequiv ac'|
%   then |(ap finv ac) clequiv (ap finv ac')|) and
%   |.->| (if
%   |cc creduce cc'| then |(ap f cc) areduce (ap f cc')| and
%   if |ac areduce ac'| then |(ap finv ac) creduce (ap finv ac')|)
\end{definition}

@@

This version

>>>>

Instead of using a configuration of the form
\[|iconf iS (fullconf id1 il1 tS1 ie1, fullconf id2 il2 tS2 ie2 ldots)|\]
as is done in |specLangML alpha|
where every task has its
own ML heap, we would like a single global heap as in
|oneheapiconf iS tS (oneheapfullconf id1 il1 ie1, oneheapfullconf id2 il2 ie2, ldots)|.
To keep up a fictional separation between the portions
of the global heap that tasks can use, we forbid references to
cross task boundaries.  To achieve such a system, we can
adapt most of the operational rules easily to work with these
new configurations.  In the abstract system, the \textsc{I-fork}
rule created a fresh heap for the new task to execute in.  Since we
want to forbid references to cross task boundaries, we do not modify
the global heap and instead add an additional premise that the forked
expression |ie| must not contain any references.
Finally, we add the same condition to the \textsc{I-send} rule.
Tasks that try to fork or send an expression that contains a
reference will now get stuck, and will stop executing by the
\textsc{I-noStep} rule.
We call this family of languages |concreteLangMl alpha|, which
is still parametrized by the scheduling policy |alpha|.

@@

In |concreteLangMl alpha| we
effectively restrict the abstract semantics in some cases (namely
to not allow the passing of references) as well as using a different
way to represent the heap.  How can we be sure that these
new, \textit{concrete semantics} for our system still satisfy
non-interference?

We will use two techniques to ensure non-interference.  First,
we require some form of isomorphism to the abstract language
|specLangML alpha| that ensures that the concrete system behaves
in a similar way.  Secondly, we pose conditions on what kind
of restrictions the concrete language is allowed to make
on the abstract language.

Intuitively, there are two scenarios in which non-interference
can be violated when restricting the transition relation:
\begin{itemize}
  \item Adding an additional premise to one of the rules that inspects
  information which is above the security label of the task currently
  executing.  This would allow another task to observe the termination
  of the other if both are at the same security level, and thus observe
  the secret information that was used in the additional
  premise of the rule.
  \item Adding additional restrictions to transition rules can
  also cause the whole configuration to get stuck, which can leak
  information through the termination channel (which is only
  of interest for TSNI, but not TINI, of course).
\end{itemize}

Next, we will make these two intuitions precise and give a general
definition for how non-interference can be preserved.  Then, we
show that the language |concreteLangMl alpha| satisfies these
conditions.

\Red{Should we give an example why removing transitions in general
  is not OK?  Edward has already typeset one that we could
  use.}

@@

Now, we can propose the following two-step process to ensure that
non-interference is preserved:
\Red{We should find a better name for this, or not use a name, or something.}
\begin{definition}[Non-interference preserving, restricted isomorphism (NPRI)]
  We call an IFC language
  |concreteLang alpha targetLang| (the concrete language for a given
  target language |targetLang| and scheduling policy |alpha|)
  a \textit{non-intereference preserving,
  restricted isomorphism} with regard to the IFC language
  |specLang alpha targetLang|
  (the abstract language) if the following
  two conditions hold:
  \begin{enumerate}
    \item Any of the reduction rules of the abstract
    semantics |specLang alpha targetLang|
    presented in Section~\ref{sec:retrofit} can
    be restricted by adding a predicate $P$ to the premise of
    any rule other than \textsc{I-noStep}.  The predicate $P$
    can depend only on the \textit{erased} configuration
    |erase il ic|, where |il| is the label of the first task
    in the task list and |ic| the full configuration.
    \item The concrete language
    |concreteLang alpha targetLang|
    must be isomorphic to the abstract semantics |specLang alpha targetLang|
    with the additional predicates $P$ on the rules.
  \end{enumerate}
\end{definition}

\begin{theorem}[NPRI implies non-interference]
  \label{thm:npri}
  If a language |concreteLang alpha targetLang| is a NPRI, and |alpha|
  is a scheduling policy that makes |specLang alpha targetLang| satisfy
  TINI or TSNI, then also
  |concreteLang alpha targetLang| satisfies TINI, or TSNI, respectively.
\end{theorem}

We give the proof of this in Section~\ref{sec:proofs}.
Going back to our single-heap ML-like IFC language, we can
show that it is in fact NPRI, and thus non-interfering.

@@

\begin{proof}
  It is easy to see that condition (1) of NPRI is satisfied, since
  we only add a side condition to the rules \textsc{I-fork} and
  \textsc{I-send}, and the condition only inspects the expression
  of the current task, which is by the definition of |erasef il| not
  erased.

@@

\subsection{Proof of Theorem~\ref{thm:npri}}
\label{sec:proofs}

Before we give the actual proof, we state and proof some properties
of isomorphisms.

\Red{Adapt proofs.}

\begin{lemma}
  \label{lemma:tsni}
  For any configurations |ic1|, |ic1'|, |ic2|, and label |il| where
  \begin{equation} \label{eq:tsni-lemma-lhs}
  |ic1| \approx_{|il|} |ic2|
  \qquad \text{and} \qquad
  |ic1| |.->| |ic1'|
  \end{equation}
  then there exists a configuration |ic2'| such that
  \begin{equation} \label{eq:tsni-lemma-rhs}
  |ic1'| \approx_{|il|} |ic2'|
  \qquad \text{and} \qquad
  |ic2| |.->|^* |ic2'|
  \ \text{.}
  \end{equation}
\end{lemma}

\begin{proof}
  We proof the theorem by induction on the length of the derivation sequence in~\eqref{eq:tsni-lhs}.
  The base case for derivations
  of length 0 is trivial, allowing
  us to simple chose $|ic2'=ic2|$.  In the step case, we assume
  the theorem holds for derivation sequences of length up to $n$, and show that it also
  holds for those of length $n+1$.  We split the derivation sequence from~\eqref{eq:tsni-lhs} as follows:
  \[
  |ic1| |.->| |ic1''| |.->|^n |ic1'|
  \]
  for some configuration |ic1''|.  By Lemma~\ref{lemma:tsni}, we get
  |ic''| with
  \begin{equation} \label{eq:tsni-proof-1}
  |ic1''| \approx_{|il|} |ic2''|
  \qquad \text{and} \qquad
  |ic2| |.->|^* |ic2''|
  \end{equation}
  Applying the induction hypothesis to
  $|ic1''| |.->|^n |ic1'|$, we get |ic2'| with
  \begin{equation} \label{eq:tsni-proof-2}
  |ic1'| \approx_{|il|} |ic2'|
  \qquad \text{and} \qquad
  |ic2''| |.->|^* |ic2'|
  \end{equation}
  Stitching together the derivation sequences from~\eqref{eq:tsni-proof-1} and~\eqref{eq:tsni-proof-2} directly gives
  us~\eqref{eq:tsni-rhs}, which concludes the proof.
\end{proof}
\begin{proof}[Proof of Lemma~\ref{lemma:tsni}]
  We proof the lemma by induction on the derivation
  for~\eqref{eq:tsni-lemma-lhs}, and consider all possible rules that
  could have been used to make the reduction step.
  First, we observe there must be at least one task in |ic1|, otherwise
  it could not take a step.  Thus, |ic1| is of the form
  |iconf iS1 (it1, its1)|.  Consider two cases:
  \begin{itemize}
    \item $|erase il it1|=|bullet|$. In this case, we do not need to take a
    step for
    |ic2|, because |ic2'=ic2| will already be |il|-equivalent to |ic1'|.
    To see that, note that the tasks |its1| in |ic1| are left in the
    same order and unmodified. The task |it1| either
    gets dropped (by \textsc{I-noStep}), or
    transforms into a new task |it1'| as well as potentially spawning a new
    task |it1''|.  Since both |it1'| and |it1''| have a label that is
    at least as high as the label of |it1|, they will get filtered
    by |erasef il| in |ic1'|.
    Lets consider the possible changes to |iS1|: If |it1| sends a
    message, then the corresponding queue in |iS1'| will contain one
    more message.  However, that message has the label of |it1| (or a higher
    label), and will thus get erased by |erasef il|.
    Finally, receiving or not receiving a message (by
    \textsc{I-recv} or \textsc{I-noRecv}) only changes the queue of
    task |it1|, which are erased from |iS1|.
    This ensures that $|ic1'|\approx_{|il|}|ic2'|$.

    \alphacondition{We need all scheduling policies to not change the order
      of any tasks (except for the first one).  Newly spawned task can appear
      anywhere in the list.}
    \item $|erase il it1|\neq|bullet|$.  Here, there must be a corresponding
    task |it2| in |ic2|,
    such that |it1=it2| (otherwise |ic1| and
    |ic2| could not be |il| equivalent).
    However, |it2| might not be at the beginning of the task list yet, but
    all tasks occurring before it must get erased by |erasef il|.
    In |ic2|, we can first take some number of steps until |it2| moves
    to the front of the list (\Red{Say more precisely why
      (basically because we require the scheduling policy to do that)}).

    \alphacondition{The scheduling policy must eventually let any task in
      the task list evaluate.  In particular, it cannot get stuck when the
      first task gets stuck, or keep executing a small number of tasks
      exclusively forever (e.g. just execute the first task all the time
      if it gets into an infinite loop).}

    Therefore, after |ic2| has potentially executed some number of steps
    to arrive at |ic2''|, we are now in the situation where $|ic1|\approx_{|il|}|ic2''|$, and the first task |it1| and |it2| don't
    get erased and are thus equivalent.
    The task |it2| can now take exactly the same step as |it1|;  we only
    need to argue that the potential differences in |iS1| and |iS2| cannot
    have an influence on the execution (and we know |iS1| and |iS2| are
    |il| equivalent).
    Again, only send and receive will change |iS|, so we consider all
    rules that could have been used for~\eqref{eq:tsni-lemma-rhs}.  If
    a message is sent, then the message is added to the same queue.  This
    queue is either completely erased, or it is |il| equivalent.  In both
    cases, |il| equivalence of |iS1'| and |iS2'| is preserved.
    When the tasks are receiving a message, then by the reduction rules
    we know that they first filter the queue by the label of |it1|.  We
    also know that the queues are equivalent when filtered by the less
    restrictive label |il|, thus the messages received (or dropped) from the
    queue are equivalent.
  \end{itemize}
\end{proof}

@@

@@
